Better fusion for filters

### 2014-03-28 2014betterfusionforfilters/section/04-ILP.tex
minor fix
simplified ILP constraints

\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> $-N x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \>             (no edge from i to j)            \\
           \>    $x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \>             (an edge from i to j)            \\
\\
           \> $x_{ij}$    \> $\le$ \> $c_i$           \>       \>            \\
           \> (for all edges from i to j; thus $c_i=0$ only if all $x{ij} = 0$)      \\
\\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (fusion-preventing edges from i to j)      \\
\\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_i=\emptyset \wedge gen_j=\emptyset$)                 \\
           \> $x_{i'j}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_i=\{i'\}$) \\
           \> $x_{ij'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_j=\{j'\}$) \\
\end{tabbing}
It is worth noting that for the edge cases, $x_{ij} = 1 \implies \pi_i < \pi_j$, meaning if $i$ and $j$ are not fused together, $j$ must be scheduled after $i$.

@@

\subsection{Subtype optimisations}
For two clusters $i, j$, if $x_{ij} = 0 \vee \tau_i \not= \tau_j$, then $i$'s generator must be fused with $j$ and vice versa, recursively ($x_{ij} = 0 \implies check(i,j)$).
Problem: actually generating constraints for \emph{all} pairs of nodes with unequal types is excessive.
I see two solutions to this:
\begin{itemize}
\item
Generate fewer constraints and rely on transitivity.
This means that for each pair $i,j$ we wouldn't generate an $x_{ij}$ and thus not generate those constraints, unless $i$ or $j$ is a @filter@, or arc between $i$ and $j$.
This would help, but also means we can't \emph{minimise} on these removed $x_{ij}$, so the solution won't count the (rather smaller) benefits of fusing two non-connected nodes.
\item
Alternatively:
we can say, only generate subtype constraints if there is no \emph{fusion preventing path} between $i$ and $j$.
Here, we generate fewer constraints than originally, but still more than above.
We also must preprocess the graph to find such blocking paths.
But we still get the minimisation to count non-connected nodes.
\end{itemize}

>>>>

\subsubsection{The simple version}
\TODO{Explain this simple version}
\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> $-N x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \>             (no edge from i to j, but there is a fusion benefit)            \\
           \>    $x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \>             (an edge from i to j)            \\
\\
           \> $x_{ij}$    \> $\le$ \> $c_i$           \>       \>            \\
           \> (for all edges from i; thus $c_i=0$ only if $\forall j | (i,j) \in E. x_{ij} = 0$)      \\
\\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (fusion-preventing edges from i to j)      \\
\\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_i=\emptyset \wedge gen_j=\emptyset$)                 \\
           \> $x_{i'j}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_i=\{i'\}$) \\
           \> $x_{ij'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_j=\{j'\}$) \\
\end{tabbing}
It is worth noting that for the edge cases, $x_{ij} = 1 \implies \pi_i < \pi_j$, meaning if $i$ and $j$ are not fused together, $j$ must be scheduled after $i$.

\subsubsection{A slightly smarter version}
We can simplify the version above and remove many constraints and variables by noting that if there is a fusion-preventing edge between $i$ and $j$, $x_{ij} = 1$, so the precedence constraint can be simplified to just $\pi_i < \pi_j$, and the $x_{ij}$ variable removed.
Likewise with clusters of different, incompatible types.
The contraction variable $c_i$ can be removed, if $i$ has any fusion-preventing output edges: any such edges use the array and cannot be fused, thus there is no possibility of the array being contracted away.

\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> $-N x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \>             (no edge from i to j, but there is a fusion benefit)            \\
           \>    $x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \>             (an edge from i to j)            \\
           \>             \>       \> $\pi_i < \pi_j$ \>       \>            \\
           \>             (a \emph{fusion-preventing} edge from i to j)            \\
\\
           \> $x_{ij}$    \> $\le$ \> $c_i$           \>       \>            \\
           \> (for all edges from i with no fusion-preventing outputs)      \\
\\
           \> $x_{i'j}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_i=\{i'\}$) \\
           \> $x_{ij'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_j=\{j'\}$) \\
\end{tabbing}

@@

\subsection{Subtype optimisations}
For two clusters $i, j$, if $x_{ij} = 0 \vee \tau_i \not= \tau_j$, then $i$'s generator must be fused with $j$ and vice versa, recursively ($x_{ij} = 0 \implies check(i,j)$).
Problem: actually generating constraints for \emph{all} pairs of nodes with unequal types is excessive.
I see two solutions to this:
\begin{itemize}
\item
Generate fewer constraints and rely on transitivity.
This means that for each pair $i,j$ we wouldn't generate an $x_{ij}$ and thus not generate those constraints, unless $i$ or $j$ is a @filter@, or arc between $i$ and $j$.
This would help, but also means we can't \emph{minimise} on these removed $x_{ij}$, so the solution won't count the (rather smaller) benefits of fusing two non-connected nodes.
\item
Alternatively:
we can say, only generate constraints if there is no \emph{fusion preventing path} between $i$ and $j$.
Here, we generate fewer constraints than originally, but still more than above.
We also must preprocess the graph to find such blocking paths.
But we still get the minimisation to count non-connected nodes.
\end{itemize}
